Forecasting the 2024 U.S. Presidential Election: A Poll-of-Polls Approach for Predicting the Outcome

### 2024-11-04 paper/paper.qmd
diagram1

library(tidyverse)
library(ggplot2)
library(MASS)
library(dplyr)
```

@@

# Description of the Model

@@

```{r}
# Load necessary libraries
library(rstanarm)
library(tidyverse)

# Load the cleaned data
polls_data_cleaned <- read_csv("../data/02-analysis_data/cleaned_president_polls.csv")

# Fit the logistic regression model
win_model <- stan_glm(
  formula = win ~ pollscore + sample_size + party_binary + pct,
  data = polls_data_cleaned,
  family = binomial(link = "logit"),
  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
  seed = 853
)

# Print summary of the model
summary(win_model)
```
## Model Performance and Interpretation
The model's output includes estimates for each predictor, providing insights into their relative importance in predicting a candidate's probability of winning. The summary of the model reveals the following key points:

@@

```{r}
# Calculate predicted win probabilities
polls_data_cleaned <- polls_data_cleaned %>%
  mutate(predicted_win_prob = predict(win_model, type = "response"))

@@

# Save the model and predictions for later use
saveRDS(win_model, "../models/election_win_model.rds")
write_csv(polls_data_cleaned, "../data/04-model_results/polls_with_predictions.csv")
```
The model described above allows us to predict the likelihood of each candidate winning based on the given polling data, using a logistic regression approach. The model was implemented using rstanarm to provide a Bayesian estimation, which helps incorporate uncertainty into our predictions.

>>>>

library(tidyverse)
library(ggplot2)
library(MASS)
library(dplyr)
library(rstanarm)

@@

## Description of the Model

@@

```{r}
#| echo: false
# Load necessary libraries
library(rstanarm)
library(tidyverse)

# Load the cleaned data
polls_data_cleaned <- read_csv("../data/02-analysis_data/cleaned_president_polls.csv")
# Fit the logistic regression model
options(mc.cores = parallel::detectCores(), rstanarm.verbosity = 0)

win_model <- stan_glm(
  formula = win ~ pollscore + sample_size + party_binary + pct,
  data = polls_data_cleaned,
  family = binomial(link = "logit"),
  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
  seed = 853,
  refresh = 0  # Suppress progress output during sampling
)
```

```{r}
#| echo: false
# Extract summary of the model without sampling details
summary_text <- capture.output(print(summary(win_model)))
cat(summary_text, sep = "\n")

## Model Performance and Interpretation

The model's output includes estimates for each predictor, providing insights into their relative importance in predicting a candidate's probability of winning. The summary of the model reveals the following key points:

@@

```{r}
#| echo: false

# Calculate predicted win probabilities
polls_data_cleaned <- polls_data_cleaned %>%
  mutate(predicted_win_prob = predict(win_model, type = "response"))

@@

# Save the model and predictions for later use
saveRDS(win_model, "../models/election_win_model.rds")
write_csv(polls_data_cleaned, "../data/04-model_results/polls_with_predictions.csv")
```

The model described above allows us to predict the likelihood of each candidate winning based on the given polling data, using a logistic regression approach. The model was implemented using rstanarm to provide a Bayesian estimation, which helps incorporate uncertainty into our predictions.

```{r}
#| echo: false
# Plot relationship between poll percentage and predicted win probability
polls_data_cleaned %>%
  ggplot(aes(x = pct, y = predicted_win_prob)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "loess", se = FALSE, color = "red") +
  labs(title = "Relationship Between Poll Percentage and Predicted Win Probability",
       x = "Poll Percentage",
       y = "Predicted Win Probability") +
  theme_minimal()
```

```{r}
#| echo: false
# Plot predicted win probabilities by party
polls_data_cleaned %>%
  ggplot(aes(x = party_binary, y = predicted_win_prob, fill = factor(party_binary))) +
  geom_boxplot() +
  scale_fill_manual(values = c("0" = "red", "1" = "blue"), labels = c("0" = "Republican", "1" = "Democratic")) +
  labs(title = "Predicted Win Probabilities by Party",
       x = "Party",
       y = "Predicted Win Probability") +
  theme_minimal()
```

```{r}
#| echo: false
# Plot predicted win probability distribution
ggplot(polls_data_cleaned, aes(x = predicted_win_prob)) +
  geom_histogram(binwidth = 0.05, fill = "steelblue", color = "black") +
  labs(title = "Distribution of Predicted Win Probabilities",
       x = "Predicted Win Probability",
       y = "Count") +
  theme_minimal()
```
